Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2019-02-13 WholePlotTrajectories.Rmd
intro

An early conceptual basis of the linkage between biodiversity and disturbance is the Intermediate Disturbance Hypothesis (IDH).
The IDH assumes a relationship between community diversity and the intensity and frequency of disturbance events [@Connell1978].
The theory and states that community diversity peaks at intermediate level of disturbance.
The fluctuations of environmental conditions after disturbance are assumed to prevent competitive exclusion in fostering both competitively superior species and fast colonizers [@Shea2004;@Pulsford2016].
In tropical forests, however, observations of the IDH are often blurry compared to theoretical expectations [@Hugues2007;@Sheil2003;@Norden2017], and the validation of the theory remains controversial [@Hubbell2001;@Fox2013;@Sheil2013].
The IDH may result from several processes and in the highly diverse tree communities its pattern might differ from those recognized otherwise.
Most of the studies rely on a single diversity metric, often the taxonomic richness [@Molino2001], which is limited to grasp all facets of the diversity-disturbance relationship.
In this study we aimed as well to clarify the scope of the IDH and discuss its implication for the fate of tropical forests [@Sheil2003;@Shea2004].

@@

In this study, we monitored over 30 years the response of 75 ha of Neotropical forest plots set up on a gradient of disturbance intensity, from 10 to 60% of above-ground biomass (AGB) loss. We made use of a large functional traits database encompassing major leaf, stem and life-history traits in order to draw the taxonomic and functional trajectories in terms of richness, evenness, composition and redundancy. Specifically, (i) from post-disturbance trajectories we examined community taxonomic and functional response, (ii) then we discussed the scope of the IDH in taxonomic and functional facets and in time after disturbance, and (iii) we analyzed community resilience and time to recovery.

>>>>

An early conceptual basis of the linkage between biodiversity and disturbance is the Intermediate Disturbance Hypothesis (IDH).
The IDH assumes a relationship between community diversity and the intensity and frequency of disturbance events [@Connell1978].
The theory and states that community diversity peaks at intermediate level of disturbance.
The fluctuations of environmental conditions after disturbance are assumed to prevent competitive exclusion in fostering both competitively superior species and fast colonizers [@Shea2004;@Pulsford2016].
In highly diverse tropical forests, however, observations of the IDH are often blurry compared to theoretical expectations [@Hugues2007;@Sheil2003;@Norden2017], and the validation of the theory remains controversial [@Hubbell2001;@Fox2013;@Sheil2013].
The scope of the IDH should then be clarified regarding the different facets of community richness, abundance distribution and functioning.
The underlying processes of the IDH are besides part of the post-disturbance succession and the question is on the time after disturbance validating the hypothesis.
In this study we aimed as well to clarify the scope of the IDH and discuss its implication for the fate of tropical forests [@Sheil2003;@Shea2004].

@@

In this study, we monitored over 30 years the response of 75 ha of Neotropical forest plots set up on a gradient of disturbance intensity, from 10 to 60% of above-ground biomass (AGB) loss. We made use of a large functional traits database encompassing major leaf, stem and life-history traits in order to draw the taxonomic and functional trajectories in terms of richness, evenness, composition and redundancy. Specifically, (i) from post-disturbance trajectories we examined community taxonomic and functional response, (ii) then we discussed the scope of the IDH in its aspect of taxonomic and functional diversity and time after disturbance, and (iii) we analyzed community resilience and time to recovery.

### 2019-02-13 WholePlotTrajectories.tex
intro

An early conceptual basis of the linkage between biodiversity and
disturbance is the Intermediate Disturbance Hypothesis (IDH). The IDH
assumes a relationship between community diversity and the intensity and
frequency of disturbance events \citep{Connell1978}. The theory and
states that community diversity peaks at intermediate level of
disturbance. The fluctuations of environmental conditions after
disturbance are assumed to prevent competitive exclusion in fostering
both competitively superior species and fast colonizers
\citep{Shea2004, Pulsford2016}. In tropical forests, however,
observations of the IDH are often blurry compared to theoretical
expectations \citep{Hugues2007, Sheil2003, Norden2017}, and the
validation of the theory remains controversial
\citep{Hubbell2001, Fox2013, Sheil2013}. The IDH may result from several
processes and in the highly diverse tree communities its pattern might
differ from those recognized otherwise. Most of the studies rely on a
single diversity metric, often the taxonomic richness
\citep{Molino2001}, which is limited to grasp all facets of the
diversity-disturbance relationship. In this study we aimed as well to
clarify the scope of the IDH and discuss its implication for the fate of
tropical forests \citep{Sheil2003, Shea2004}.

@@

In this study, we monitored over 30 years the response of 75 ha of
Neotropical forest plots set up on a gradient of disturbance intensity,
from 10 to 60\% of above-ground biomass (AGB) loss. We made use of a
large functional traits database encompassing major leaf, stem and
life-history traits in order to draw the taxonomic and functional
trajectories in terms of richness, evenness, composition and redundancy.
Specifically, (i) from post-disturbance trajectories we examined
community taxonomic and functional response, (ii) then we discussed the
scope of the IDH in taxonomic and functional facets and in time after
disturbance, and (iii) we analyzed community resilience and time to
recovery.

>>>>

An early conceptual basis of the linkage between biodiversity and
disturbance is the Intermediate Disturbance Hypothesis (IDH). The IDH
assumes a relationship between community diversity and the intensity and
frequency of disturbance events \citep{Connell1978}. The theory and
states that community diversity peaks at intermediate level of
disturbance. The fluctuations of environmental conditions after
disturbance are assumed to prevent competitive exclusion in fostering
both competitively superior species and fast colonizers
\citep{Shea2004, Pulsford2016}. In highly diverse tropical forests,
however, observations of the IDH are often blurry compared to
theoretical expectations \citep{Hugues2007, Sheil2003, Norden2017}, and
the validation of the theory remains controversial
\citep{Hubbell2001, Fox2013, Sheil2013}. The scope of the IDH should
then be clarified regarding the different facets of community richness,
abundance distribution and functioning. The underlying processes of the
IDH are besides part of the post-disturbance succession and the question
is on the time after disturbance validating the hypothesis. In this
study we aimed as well to clarify the scope of the IDH and discuss its
implication for the fate of tropical forests
\citep{Sheil2003, Shea2004}.

@@

In this study, we monitored over 30 years the response of 75 ha of
Neotropical forest plots set up on a gradient of disturbance intensity,
from 10 to 60\% of above-ground biomass (AGB) loss. We made use of a
large functional traits database encompassing major leaf, stem and
life-history traits in order to draw the taxonomic and functional
trajectories in terms of richness, evenness, composition and redundancy.
Specifically, (i) from post-disturbance trajectories we examined
community taxonomic and functional response, (ii) then we discussed the
scope of the IDH in its aspect of taxonomic and functional diversity and
time after disturbance, and (iii) we analyzed community resilience and
time to recovery.
